---
title: "DATA607_Tidyverse Extend"
output: 
  html_document:
    theme: cerulean
    Highlight: tango
    toc: yes
    toc_float: yes
---

#### Title: CUNY SPS MDS DATA607_Tidyverse Extend"
#### Author: Charles Ugiagbe
#### Date: 11/13/2021


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Tidyverse create 

**The Part 1 Tidyverse Create Assignment was done by "Jiho Kim".**

**The Part 2: Tidyverse Extend was done by me. I added some more code to the Tidyverse create using the same dataset heart.csv**

*Source: Kagle data*


```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

### Retrieve Data using readr::read.csv

I chose heart failure dataset from Kaggle and read the csv file using read.csv function.

```{r}
heartdata <- read.csv('https://raw.githubusercontent.com/jihokim97/FALL2021TIDYVERSE/Tidyverse/heart.csv')
view(heartdata)
head(heartdata)
```

### Dplyr

I used dplyr packages to demonstrate different capabilities. 

### Filter

We can use filter functions to filter the desired rows from the dataframe.
```{r}
filteredheart <- heartdata%>%
  filter(Cholesterol > 200)
head(filteredheart)
```

### Select

We can select certain desired columns using select function.

```{r}
selectheart <- heartdata%>%
  select(c('Age','MaxHR','HeartDisease'))
head(selectheart)
```

### group_by and Summarise

We can group dataframe by columns and create new data frame into a single row using summarise.

```{r}
bysex <- heartdata%>%
  group_by(Sex)%>%
  summarise(count = n())

bysex
```



##        Part 2: Tidyverse Extend

### dplyr:: select

Q: How do i select from one column to another in a table

A: By using the dplyr select function

```{r}
heartdata2 <- heartdata%>%
  select(Age:Cholesterol,HeartDisease)
head(heartdata2)
```


### dplyr::rename

Q: How do i rename a variable in a table?

A: I use dplyr rename function

```{r}
heartdata2 <- heartdata2 %>% rename("Gender" = Sex)
head(heartdata2)
```

###   dplyr::case_when()

Q: How do i vectorise multiple if else statement

A: By Using the case_when() function

```{r}
heartdata3 <- heartdata2 %>% 
  mutate(Heart_condition = case_when(Cholesterol >= 300 ~ 'Severe'
                                     ,TRUE ~ 'Not_severe' )
         )
head(heartdata3)
```

###   `ggplot2::Barplot

Q: I want to plot a Barchart of the count of Male and Female in the dataset

A: Use the ggplot function and the geom_bar layer

```{r}
ggplot(data = heartdata2, aes(x = Gender)) + geom_bar(fill = "blue") + 
  labs(title = "Bar chart for count of Sex") + theme_bw()
```

###   coord_flip()

Q: How do I flip co-ordinates?

A: Add the coord_flip layer

```{r}
ggplot(data = heartdata2, aes(x = Gender)) + geom_bar(fill = "blue") + 
  labs(title = "Bar chart for count of Sex") + theme_bw() + coord_flip()
```


###    ggplot: Histogram

```{r}
# Histogram of Cholesterol level
ggplot(heartdata2, aes(x = Cholesterol)) + geom_histogram(binwidth = 30, fill = "purple") + 
  labs(title = "Distribution of Cholesterol level") + ylab(NULL)
```

