---
title: "working with Json Extra Credit"
author: "Charles Ugiagbe"
date: "10/20/2021"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---

### Title: CUNY SPS MDS Data606_working with Json"
### Author: Charles Ugiagbe


```{r message=FALSE, warning=FALSE}
library("httr")
library("ggplot2")
library("jsonlite")
library("plyr")
library("dplyr")
library("knitr")
```

## Getting the data

We use the [Nobel Prize API](https://nobelprize.readme.io/) to fetch laureate data in JSON format.

Purpose: This data seek to analyse the different parameter in the Nobel Prize API above. This capture Nobel Prize winner from 1975

```{r get-data, tidy=TRUE, cache=FALSE, message=FALSE}
url <- "http://api.nobelprize.org/v1/laureate.json"
nobelprz <- fromJSON(url)
names(nobelprz)
```


```{r}
names(nobelprz$laureates)
```

```{r}
names(nobelprz$laureates$prizes[[1]])
```



## Nobel Prizes won by Gender and organization

```{r}
gender <- as.data.frame(table(nobelprz$laureates$gender), stringsAsFactors = FALSE)
ggplot(gender) + geom_bar(aes(Var1, Freq), stat = "identity", fill = "skyblue3") + theme_bw() + labs(x = "gender", y = "count", title = "All Nobel Prizes by Gender")
```
##    Comparing the Nobel prizes won by Gender and their various categories

More male has won the nobel prize than female in all categories

```{r}
cnt <- sapply(nobelprz$laureates$prizes, function(x) nrow(x))
prizes <- ldply(nobelprz$laureates$prizes, as.data.frame)
prizes$id <- rep(nobelprz$laureates$id, cnt)
prizes$gender <- rep(nobelprz$laureates$gender, cnt)
pg <- as.data.frame(table(prizes$category, prizes$gender), stringsAsFactors = FALSE)
ggplot(pg, aes(Var2, Freq)) + geom_bar(aes(fill = Var2), stat = "identity") + theme_bw() + facet_grid(. ~ Var1) + labs(x = "gender", y = "count", title = "All Nobel Prizes by Gender and Category") + scale_fill_manual(values = c("darkorange", "skyblue3", "grey"), name = NULL) + theme(axis.text.x = element_blank())
```


##   Multiple winners of the Nobel Prize - individual and Organization

```{r}
multi <- which(sapply(nobelprz$laureates$prizes, function(x) nrow(x)) > 1)
winners <- nobelprz$laureates[multi, c("firstname", "surname", "born", "bornCountry")]
print(kable(winners), type = "html", comment = FALSE, include.rownames = FALSE)
```

## Share Nobel Prize won by two and Three Laureates


```{r}
share <- as.data.frame(table(prizes$year, prizes$category), stringsAsFactors = FALSE)
colnames(share) <- c("year", "category", "Freq")
share1 <- as.data.frame(table(share$Freq), stringsAsFactors = FALSE)
ggplot(share1[share1$Var1 != 0, ]) + geom_bar(aes(Var1, Freq), stat = "identity", fill = "skyblue3") + theme_bw() + labs(x = "number of laureates", y = "count", title = "Laureates per Nobel Prize")
```

##  Age Distribution of  Nobel laureates 

```{r}
prizes$born <- rep(nobelprz$laureates$born, cnt)
prizes$age <- as.Date(paste(prizes$year, "12-31", sep = "-"), "%Y-%m-%d") - as.Date(prizes$born, "%Y-%m-%d")
ggplot(prizes[!is.na(prizes$category), ]) + geom_boxplot(aes(category, as.numeric(age) / 365), fill = "red") + theme_bw() + stat_summary(aes(category, as.numeric(age) / 365), fun.y = "median", geom = "point") + labs(x = "category", y = "age (years)", title = "Age Distribution of Nobel Laureates by Category")
```
```

