---
title: "Web APIs"
output: 
  html_document:
    theme: cerulean
---

```{r global_options, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Title: CUNY SPS MDS DATA607_WK9Assignmt"
### Author: Charles Ugiagbe
### Date: "10/24/2021"


<style>
div.aquamarine { background-color:#d9eff9; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

**Assignment Question**

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis(https://developer.nytimes.com/apis)

You’ll need to start by signing up for an API key.

Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

**Approach to Solution**

The New York Times API that interest me the most was “Books API”. After requesting the API key, I was able to access it in R. I extracted the data of “hardcover fiction books” from the books API, and this was in JSON format. I then used “fromjson” to convert the JSON data to R objects.

</div> \hfill\break



##        Load the Required Packages

```{r message=FALSE, warning=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(kableExtra)
```

##        Read the API
```{r}
bookAPI <- "https://api.nytimes.com/svc/books/v3/lists/current/hardcover-fiction.json?api-key=yrHCGK1wjdTA8abjGAyni4Q1A4RiWACs"
```


### Transform Json Data into R Dataframe and take a glimpse view of the data structure

```{r}
fiction_books <- fromJSON(bookAPI)[[5]][[11]]
glimpse(fiction_books)
```



### Create a new Dataframe from the Original one for Analysis.

```{r}
fiction_books2 <- fiction_books[c("rank", "publisher", "title", "author", "contributor", "primary_isbn13")]

fiction_books2 %>%
  kbl(caption = "Hardcover fiction books") %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "Red")
```


###       Extra Analysis

```{r}
# Publisher df. Order books by publisher

df_pub <- fiction_books2 %>%
  group_by(publisher) %>%
  summarise(books_published = n())
```


**Plot the ranked books by publisher and visualize which publicher has more books in the ranking.**

```{r}
df_pub <- df_pub[order(-df_pub$books_published), ]

df_pub %>%
  
  ggplot(aes(reorder(publisher, books_published), books_published)) +
  
  geom_col(aes(fill = books_published)) + 
  
   scale_fill_gradient2(low = "Green",
                       high = "Blue",
                       ) +
  
   labs(title="Ranked hardcover fiction books by publisher") + ylab("books_published") + theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x=element_text(angle=45)) + theme(plot.title = element_text(hjust=0.5)) + theme(axis.text.x = element_text(margin = margin(t = 25, r = 20, b = 0, l = 0))) 
  
  
  
```

