---
title: "Tidying and Transforming Wide Data"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---



### Title: CUNY SPS MDS DATA607_PROJ2"
### Author: Charles Ugiagbe
### Date: "10/3/2021"



##    Untidy Data 1:  Submitted by Eric Lehmphul

```{r}
library(tidyverse)
```


###    Read the data
```{r}
url <- "https://raw.githubusercontent.com/omocharly/DATA607_PROJECTS/main/Untidy%20Data%20-%20By%20Eric%20Lehmphul.csv"
testdata <- read.csv(url)
testdata
```

###  Tidy the data by separating the student score fromt the time.
```{r}
tidy_testdata = testdata %>%
  separate(Test1..TimeStudiedTest1,
           into = c("Test1","TimeStudiedTest1"),
           sep = ", ", convert=TRUE) %>%
  separate(Test2..TimeStudiedTest2,
           into = c("Test2","TimeStudiedTest2"),
           sep = ",", convert=TRUE) %>%
  separate(Test3..TimeStudiedTest3,
           into = c("Test3","TimeStudiedTest3"),
           sep = ", ", convert=TRUE) %>%
  separate(Test4..TimeStudiedTest4,
           into = c("Test4","TimeStudiedTest4"),
           sep = ", ", convert=TRUE)

# To rearrnge the Columns
test_data <- tidy_testdata %>% select(Student, Gender, Test1, Test2, Test3, TimeStudiedTest1, TimeStudiedTest2, TimeStudiedTest3)
test_data
```

Replace "NA" with Zero(0).
```{r}
test_data1 <- test_data %>% replace(is.na(.), 0)
test_data1
```

###       Calculate the Average scores and Time studied.
```{r}
test_score <- mutate(test_data1,
  Avg_score = round((Test1 + Test2 + Test3)/3,0),
  Avg_studytime = round((TimeStudiedTest1 + TimeStudiedTest2 + TimeStudiedTest3)/3,0))
test_score
```

### Select the needed Column
```{r}
test_score <- select(test_score, Student, Gender, Avg_score, Avg_studytime)
test_score
```

### Observe the Plot of Average score against studied Time.

```{r}
scorevtime <- ggplot(data = test_score, aes(Avg_studytime, Avg_score)) + geom_point() +
  geom_smooth(se = FALSE) + ylab("Average Score") + xlab("Average Study Time") + theme_bw() + 
  labs(title = "Average Test Score vs Average Study Time")
scorevtime
```


**The correlation between them is  0.7278**

```{r}
x <- test_score$Avg_studytime
y <- test_score$Avg_score
cor.test(x, y)
```


--------------------------------------------------------------------------------

\clearpage




##    Untidy Data 2: Submitted by me (Charles Ugiagbe)

###     Read the Data

```{r}
url2 <- "https://raw.githubusercontent.com/omocharly/DATA607_PROJECTS/main/Untidy%20Data%20%20-%20Charles%20Ugiagbe.csv"
Exam_score <- read.csv(url2)
head(Exam_score)
```

### Clean the Data; Extract the Date from the Timestamp

```{r}
Exam_score$date_component <- as.Date(Exam_score$Timestamp,'%d/%m/%Y')
glimpse(Exam_score)
```


###   Select the Column of interest

```{r}
Exam_score <- select(Exam_score, date_component, Entering.Grade.Level, Birth.Month, OLSAT.Verbal.Score, NNAT.Non.Verbal.Raw.Score)
head(Exam_score)
```     


```{r}
Exam_score1 = Exam_score %>% rename(
      Date = date_component,
      Entering_Grade_lvl = Entering.Grade.Level,
      Birth_month = Birth.Month,
      Verbal_score = OLSAT.Verbal.Score,
      Non_verbal_score = NNAT.Non.Verbal.Raw.Score)
Exam_score1
```


###   Take the Plot of Verbal Vs Non-Verbal

```{r}
Exam_score2 <- ggplot(data = Exam_score1, aes(as.numeric(Verbal_score), as.numeric(Non_verbal_score))) + geom_point() +
  geom_smooth(se = FALSE) + ylab("Verbal Score") + xlab("Non Verbal Score") + theme_bw() + 
  labs(title = "Verbal Score vs Non Verbal Score")
Exam_score2
```



--------------------------------------------------------------------------------

\clearpage



##     Untidy Data 3: Submitted by Mauricio Claudio

```{r}
url3 <- "https://raw.githubusercontent.com/omocharly/DATA607_PROJECTS/main/Untidy%20Data%20-%20by%20Mauricio%20Claudio.csv"
churn_data <- read.csv(url3)
churn_data
```

###  Clean and Tidy the Data
```{r}
churn_data[churn_data==""]=NA
churn_new = churn_data %>%
  select(-Description) %>%
  fill("Division") %>%
  pivot_longer(c(Jan,Feb,Mar,Apr,May,Jun,Jul,Aug,Sep,Oct,Nov,Dec),
  names_to="Month",
  values_to = "Customers",
  names_transform = list(Customers=as.integer)
  )

churn_new
```


###   To test the most Effective Division using the ratio by Substracting loss from gain
```{r}
churn_new %>% group_by(Division) %>%
  summarize(Net_Customers=sum(Customers)) %>%
  ggplot(aes(Division,Net_Customers)) + geom_col()
```

###    To show which month Generate the Most Customers
```{r}
churn_new %>%
  group_by(Month) %>%
  summarize(Net_Customers=sum(Customers)) %>%
  ggplot(aes(Month,Net_Customers)) + geom_col()
```




